### 2020-08-03 Scripts/Comparing_climate_simulations.Rmd

---
title: "A simulated scenario"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)

```

## The nested-for-loop scenario

```{r}
data_for_loop_1 <- read_csv("Output/simulation.from.real.for.loop.csv")

data_for_loop_2 <- read_csv("Output/simulation.from.std.for.loop.csv")

data_from_function <- read_csv("Output/simulation.from.std.functions.csv")

data_from_function_cleaned <- read_csv("Output/simulation.from.std.functions.cleaned.csv")

data_from_function_cleaned %>% 
  select(Area, Year, Temperature = TemperatureCelsius, pH = pH_units) -> data_from_function_cleaned
```

```{r}
now.env.data <- read_csv(here("Input","env.data.for.log.regression.csv")) %>% 
  mutate (Year = 2017) %>% 
      filter(pH > 7) %>%
      mutate(
        Season = case_when(
          str_detect(sample, "1703|1710|1711|1801|1803") ~ "Winter",
          TRUE ~ "Summer"
        ),
        Area = case_when(
          str_detect(sample, "CP|LK|FH") ~ "SJI",
          TRUE ~ "Hood Canal"
        )
      )
```

```{r}

list.of.dfs <- list(data_for_loop_1,data_for_loop_2,data_from_function, data_from_function_cleaned, now.env.data )

names(list.of.dfs) <- c("for loop from real", "for loop from std" , "function data", "function data cleaned", "real data")

all.together <- bind_rows(list.of.dfs, .id = "Source")

```

Let's see them in hex
```{r}
all.together %>% 
  filter (Area == "SJI", Source != "for loop from real") %>% 
  ggplot(aes(x = Temperature, y = pH)) +
  geom_hex(bins= 40, aes(group = as.factor(Year), fill = ..density..)) +
  facet_grid (Year~Source) +
  ggtitle("San Juan Islands") -> left

all.together %>% 
  filter (Area != "SJI", Source != "for loop from real") %>% 
  ggplot(aes(x = Temperature, y = pH)) +
  geom_hex(bins= 40, aes(group = as.factor(Year), fill = ..density..)) +
  facet_grid (Year~Source) +
  ggtitle("Hood Canal") -> right

library(patchwork)

(left + right)  +  plot_layout(guides = "collect")-> all
all & scale_fill_continuous(limits = c(0,0.1))
```

```{r Check each }

expl.plots <- function (tibble){

tibble %>% 
  separate(sample, into = c("Site", "Date")) %>% 
  mutate(Date = lubridate::ymd(paste0(Date, "01")) %>% month(label = T)) ->temp

temp %>% 
ggplot(aes(x = Date, y = Temperature, fill = as.factor(Year))) +
  #geom_point(position = "jitter")+
  geom_boxplot()+
#  geom_smooth(aes(x = as.numeric(Date)), method = "loess") +
  facet_wrap(~Area)+ scale_fill_discrete(name = "Year", direction = -1) +
  guides(fill = "none") +
  labs(x = "Month") -> l1

temp %>% 
ggplot(aes(x = Date, y = pH, fill = as.factor(Year))) +
  #geom_point(position = "jitter")+
  geom_boxplot()+
#  geom_smooth(aes(x = as.numeric(Date)), method = "loess") +
  facet_wrap(~Area)+ scale_fill_discrete(name = "Year", direction = -1) +
  guides(fill = "none") +
  labs(x = "Month") -> l2

temp %>% 
ggplot(aes(x = Date, y = Temperature, color = as.factor(Year))) +
  #geom_point(position = "jitter")+
 # geom_boxplot()+
  geom_smooth(aes(x = as.numeric(Date)), method = "loess") +
  facet_wrap(~Area)+ scale_color_discrete(name = "Year", direction = -1) +
  labs(x = "Month") -> l3

temp %>% 
ggplot(aes(x = Date, y = pH, color = as.factor(Year))) +
  #geom_point(position = "jitter")+
  #geom_boxplot()+
  geom_smooth(aes(x = as.numeric(Date)), method = "loess") +
  facet_wrap(~Area)+ scale_color_discrete(name = "Year", direction = -1) +
  labs(x = "Month") -> l4

library(patchwork)

(l1+l2)/(l3 +l4) + patchwork::plot_layout(guides = "collect") 
}

```

```{r}
expl.plots(data_for_loop_1)
```

```{r}
now.env.data %>% ggplot(aes(x = pH , y = Temperature)) + geom_smooth(method = "lm") + facet_wrap(~Area) + geom_point ()

```