### 2020-08-13 Scripts/Prepare.data.for.stan.Rmd

---
title: "Prepare data for stan"
output: html_notebook
---

Turn the data into the format needed for stan.

```{r}
library(tidyverse)

plankton <- read_csv(here("Input","Combined_Biol_Env_Plankton.csv"))
phylum.info <- read_csv(here("Input","higher_taxonomy.csv")) %>% 
  unite(family, genus, species, sep = "|", col = "taxa") %>% 
  select(phylum, taxa)

plankton %>% 
  mutate(sample = str_remove_all(event, "-"),
         sample = str_remove(sample, "01$")) %>% 
  select(Area, sample, taxa, DIC = mean.DIC, Salinity = Salinity.new, pH = pH_new, Temperature) %>% 
  mutate(binary = 1) %>% 
  pivot_wider(names_from = taxa, values_from = binary, values_fill = list(binary = 0)) %>% 
  pivot_longer(cols = 7:236, names_to = "taxa", values_to="binary") %>% 
  filter(sample != "LL_201711") %>% 
  filter(!str_detect(taxa, "Gallus")) %>%
  left_join(phylum.info) %>% write_csv(here("Input","data.for.stan.csv"))

```