### 2020-09-24 Scripts/insect.all.Rmd

---
title: "Growing the database and annotate sequences"
output: html_notebook
---

This notebook contains all the steps taken to taxonomically annotate the sequences found in the Hood Canal. We will start by loading the hash keys from all sequencing runs and create a fasta file with all unique sequences

```{r load libraries, echo = FALSE}

library (tidyverse)
library (insect)
library (seqinr)
library (googlesheets)
library (here)
```

```{r load objects by the end of the cleaning }

Hash     <- read_csv (here("Output", "Hash_Key_all_together.csv")) %>% 
  select(Hash, Sequence) %>% distinct()
ALL.ASVs <- read_csv (here("Output","ASV_table_all_together.csv"))
tree.2   <- read_rds (here("Input", "classifier.rds"))

# we will also load the previous classification so we don't do the same thing all over again

previous.effort <- read_csv(here("Input", "insect.2019.csv"))
```

So these sequences are in the same direction, and they have some shared and some new sequences.

Let's keep only the unique sequences

That's 4,849 unique sequences. THis is going to take a while

Make them into a DNAbin object for insect

```{r make it into a DNA object}

new.set <- anti_join(Hash, previous.effort, by = c("Hash" = "representative"))

all.hashes.insect <- char2dna(new.set$Sequence)

names (all.hashes.insect) <- new.set$Hash

all.hashes.insect

```

Now, we will load v3 of the metazoan classification tree from insect v1.1.1

```{r taxonomy import}

#tree <- readRDS ("/Users/ramongallego/GoogleDrive/Collaborations/insect/metazoan_COI_marine_v3/classification_tree.rds")

```

And finally, classify all sequences in our bin file

```{r classify}

clasif.hashes <- classify (x = all.hashes.insect, tree = tree, cores = 4)
clasif.hases.new.tree <- classify(x = all.hashes.insect, 
                                  tree = tree.2,
                                  cores = "autodetect")

clasif.hases.new.tree %>% 
  bind_rows(previous.effort) %>% 
  write_csv(here("Input", "last.insect.csv"))

clasif.hases.new.tree %>% 
  unite (family, genus, species, sep = "|", col = "taxa")

clasif.hases.new.tree %>% dplyr::count (rank) %>% arrange(desc(n))
```
Run the translation classification

```{r classiffy translations}
INV <- getGeneticCode(id_or_name2 = "5")

Hash$AA <- as.character(suppressWarnings(translate(subseq(DNAStringSet(Hash$Sequence),start = 2),
                                                        genetic.code = INV,
                                                        no.init.codon = T)))

Hash %>% 
  mutate (stop.codons = str_count(string = AA, pattern = "\\*")) %>% 
  filter(stop.codons == 0) -> to.insect

all.hashes.insect.AA <- char2aa(to.insect$AA)

names (all.hashes.insect.AA) <- to.insect$Hash

all.hashes.insect.AA

```
And now, classify them
```{r classiffy translations2}

AAS.flatten <- lapply(to.insect$AA, char2aa)

names(AAS.flatten) <- to.insect$Hash

possibly_clasify <- possibly(~classify(x = .,
                                      tree = tree.2,
                                      cores = "autodetect",
                                      ping = 1),
                             otherwise = NA_real_)

AA.s.test <- AAS.flatten[51:60]

possibly.AAs <- map(AA.s.test, possibly_clasify)

clasif.hashes.AA <- possibly(classify(x = all.hashes.insect.AA, 
                                  tree = tree.2,
                                  cores = "autodetect", 
                             ping = 1,
                             species = "none"))

test <- all.hashes.insect.AA[19]

test.insect.AA <- classify(x= all.hashes.insect.AA, 
                           tree = tree.2,
                           ping = 1,
                           species = "none")

Hash$AA[19]
# So it breaks with stop codons
```

OK. So now let's save the classification object as an RDS

```{r save it for now}
clasif.hases.new.tree %>% 
  filter(family!= "" & phylum == "") %>% 
  distinct(class) # How many have a valid family but no phylum info

# Add new phylum info

clasif.hases.new.tree %>% 
  mutate(phylum = case_when(phylum != "" ~ phylum,
                            TRUE   ~ class))

saveRDS(clasif.hashes, paste0("hashes.annotated.",Sys.Date(),".rds"))

clasif.hashes <- readRDS("hashes.annotated.20180830.rds")

```

```{r}

thresholds <- list(0.8, 0.85, 0.95)

thresholds.classif <- map(thresholds, ~ classify(x= all.hashes.insect,
                                              tree = tree,
                                              cores = 8,
                                              threshold = .))

names(thresholds.classif) <- thresholds

#ADD THE DEFAULT

thresholds.classif[[4]] <- clasif.hashes

# Name it

names(thresholds.classif)[4] <- "0.9"

saveRDS(thresholds.classif, file =paste0("hashes.annotated.",Sys.Date(),".rds"))

list.of.thres <- readRDS(file ="hashes.annotated.2019-01-17.rds")

l2 <- lapply (list.of.thres, function (x) as.tibble(x))

```
UPDATE on NOV 6
We have added a new run of sequencing so I need to run insect on them too.

## Check the classification

And to see what is the actual resolution of our dataset, let's see how many reads can be classified to species, genus, family and order levels

```{r checking the clasiffication}

clasif.hashes %>% dplyr::count (rank) %>% arrange(desc(n))# an overview of the taxonomic resolution

All.Hashes.Classified[[1]] %>% dplyr::count (rank) %>% arrange(desc(n))

map_dfc (All.Hashes.Classified , ~ pull(., rank)) %>% mutate (representative = All.Hashes.Classified[[1]]$representative)

# a first check: How many sequences did not have a score: Either bc they exactly match a sequence in the tree, or they do not have a translation

clasif.hashes %>% 
  filter (is.na(score)) %>% # 176 Hashes
  left_join(ALL.ASVs, by = c("representative" = "Hash")) %>% 
 # group_by(representative) %>% 
  summarise(tot = sum(nReads)) # 2.55M reads - not only chimeras...

clasif.hashes %>% 
  filter(rank == "" & !is.na(score)) %>% # 2.8k  Hashes
  left_join(ALL.ASVs, by = c("representative" = "Hash")) %>% 
 # group_by(representative) %>% 
  summarise(tot = sum(nReads)) # 37M reads -  holy molly

 clasif.hashes %>% 
   filter(rank !="") %>% # 1.5 k Hashes
   left_join(ALL.ASVs, by = c("representative" = "Hash")) %>% 
 # group_by(representative) %>%
   group_by(sample) %>% 
  summarise(tot = sum(nReads)) %>% 
   ggplot(aes(x=fct_reorder(sample, tot), y = tot))+
   geom_col()# 6M reads

clasif.hashes %>% 
  filter(rank == "")

# Create the species- level dataset

map_dfc (l2, ~(dplyr::count(.,rank)))

map_dfr(l2, ~(dplyr::count(.,rank)), .id = "Threshold" ) -> Summary.thresholds.df# This puts the results in a df

Summary.thresholds.df$rank <- fct_relevel(Summary.thresholds.df$rank, "kingdom", "phylum", "subphylum", "subdivision", "superclass", "class",  "subclass", "subterclass", "infraclass", "superorder" , "order" , "suborder", "infraorder", "superfamily", "family", "subfamily" , "tribe",  "genus", "subgenus", "subsection",  "species", "subspecies", "variety")

Summary.thresholds.df$rank <- fct_recode(Summary.thresholds.df$rank, "Not Assigned" = "")

levels(Summary.thresholds.df$rank)

Summary.thresholds.df %>%
  ggplot (aes (x= rank, y = n, color = Threshold)) +
  geom_line(aes(group = Threshold))

Summary.thresholds.df %>%

  ggplot (aes (x= rank, y = n, fill = Threshold)) +
  geom_col(position = "dodge", color = "black") +
  theme(axis.text.x = element_text(angle = 45,vjust = 0.75))
  ggsave(filename = "Different.thresholds.png", width = 14)

clas.hash.tibble %>% dplyr::count (rank) %>%
  filter (str_detect(rank, ("species|genus|family|order")) ) %>%
  summarise (sum(n)) # 1331 Hashes identified to a level of order or lower

clas.hash.tibble %>% dplyr::count (rank) %>%
  filter (!str_detect(rank, ("species|genus|family|order")) ) %>%
  summarise (sum(n)) # 17k have not being IDed to that level

clas.hash.tibble %>% dplyr::count (rank) %>%
  filter ( rank == "")  %>%
  summarise (sum(n)) # 13k have a big fat 0 as a result

```
# Add the BLAST classiffication

We also classify the reads using a BLAST search, and retrieving the ???

```{r loading BLAST annotations}

all.hashes.blast     <- read_csv ("../../Analysis/Annotation/BLASTed_COI_Gallego_20190117-family.txt", col_names = c("Hash", "Family"))
all.hashes.blast.spp <- read_csv ("../../Analysis/Annotation/BLASTed_COI_Gallego_20190117-species.txt", col_names = c("Hash", "Species"))

all.hashes.blast %>% 
  left_join(all.hashes.blast.spp) -> all.hashes.blast

all.hashes.blast %>% 
  distinct(Species) # 440 unique spp

all.hashes.blast %>% 
  filter (!str_detect(Family, "idae")) %>% 
  select(-Hash) %>% 
  distinct(Family, .keep_all = T)

all.hashes.blast %>% 
  distinct(Family) # 279 unique "family"
 # get the lineages

Taxonomy.spp <- read_csv("/Users/ramongallego/GoogleDrive/Kelly_Lab/Projects/OA_eDNA/Analysis/Analysis.for.WSC/LineageLookupTable.csv") # LIneages for all spp
Taxonomy.fam <-  read_csv("/Users/ramongallego/GoogleDrive/Kelly_Lab/Projects/OA_eDNA/Analysis/Analysis.for.WSC/LineageLookupTable.families.csv")

# do we have them all?

summary (all.hashes.blast$Family %in% Taxonomy.spp$Family) # 3.3k families don't show in the lineages

all.hashes.blast %>% 
  filter (! Family %in% Taxonomy.fam$Family) %>% 
  distinct(Family) %>% 
 # mutate(match = str_which(string = Taxonomy.fam, pattern = .$Family)[1])

  mutate(Present = str_which(string = Taxonomy.fam, pattern = .$Family))#Taxonomy.fam$Phylum[str_which(string = Taxonomy.fam, pattern = .$Family)[1]],

all.hashes.blast %>% 
  left_join(clasif.hashes %>% select(Hash = representative, rank, taxon, score)) %>% 
  left_join(ALL.ASVs) %>% 
  filter (is.na(score)) %>% 
 # group_by(plant = str_detect(Family, "ceae")) %>% 
 # summarise(n.hashes = n_distinct(Hash)) # 51 are plants
  filter(!str_detect(Family, "ceae")) %>% 
  group_by(Family) %>% 
  summarise(tot = sum (nReads))

```

```{r}

all.hashes.blast %>% 
  bind_rows(new.hashes.blast) %>% 
  left_join(all.hashes.blast.spp %>%
              bind_rows(new.hashes.blast.spp), by = "Hash") %>% 
  dplyr::select (Hash,
          Blast.family = "Family",
          Blast.Species = "Species") -> BLAST.to.join

# Now extract the same info from the l2

map_dfc (All.Hashes.Classified , ~ pull(., species)) %>%
  mutate (Hash = All.Hashes.Classified[[1]]$representative) %>% 
  mutate (Insect.Spp = case_when(`0.95` != ""        ~ `0.95`,
                           `0.9`  != ""        ~ `0.9`,
                           `0.85` != ""        ~ `0.85`,
                           TRUE                ~ `0.8`)) %>%
  dplyr::select(Hash,
                Insect.Spp) -> all.hashes.insect.Spp

map_dfc (All.Hashes.Classified , ~ pull(., family)) %>%
  mutate (Hash = All.Hashes.Classified[[1]]$representative) %>% 
  mutate (Insect.Fam = case_when(`0.95` != ""        ~ `0.95`,
                           `0.9`  != ""        ~ `0.9`,
                           `0.85` != ""        ~ `0.85`,
                           TRUE                ~ `0.8`)) %>%
  dplyr::select(Hash,
                Insect.Fam) -> all.hashes.insect.Family

# Bind the two insect objects
all.hashes.insect.Spp %>% 
  left_join (all.hashes.insect.Family) %>% 
  left_join(BLAST.to.join) %>% # Now get a final classification 
  mutate(Species = case_when(Insect.Spp != "" ~ Insect.Spp,
                             TRUE ~ Blast.Species),
         Family = case_when(Insect.Fam != "" ~ Insect.Fam,
                             TRUE ~ Blast.family)) %>%
  dplyr::select(Hash, Family, Species) %>% 
  write_csv(path = paste0("Final.Classification.Hashes",Sys.Date(),".csv"))

all.hashes.insect.Family <- map_df(l2, ~ pull(.,family)) %>%
  rename_all(funs(paste0("Insect_Family_",.)))

all.hashes.insect.Species <- map_df(l2, ~pull(.,species)) %>%
  rename_all(funs(paste0("Insect_Species_",.)))

all.hashes <- bind_cols(all.hashes.insect.Family, all.hashes.insect.Species) %>%
  mutate (Hash = l2[[1]]$representative) %>%
  left_join(all.hashes.blast, .)

all.hashes.blast %>% dplyr::count(Family) # It has family and higher taxonomical info

all.hashes.blast %>% dplyr::count (Species)

```

### Now let's keep a database with five columns:

* Hash

* Sequence

* Family.insect

* Family.blast

* Species.insect

* Species.blast

And save it in disk

```{r reduced database}

clas.hash.tibble %>%
  select (Hash = representative,
          family.insect = family,
          species.insect = species) %>%
  left_join (all.hashes.blast)  -> clas.hash.tibble.to.write

write_csv(clas.hash.tibble.to.write, path = "hash.annotated.family.species.csv")

```

With this reference dataset, we were able to classify ~ 10% of the unique sequences to at least order level. 17k sequences did not have any match.

Let's chave a look at how many reads are classified

```{r combining }

ALL.ASVs %>% left_join(clas.hash.tibble.to.write, by = "Hash") -> All.ASVs.with.tax

All.ASVs.with.tax %>%
  mutate(Blast.Species = case_when(str_detect ( string = Blast.Species, pattern = "No |Not ") ~ "",
                               TRUE ~ Blast.Species) ) -> All.ASVs.with.tax

All.ASVs.with.tax %>% 
  filter(str_detect(sample, "K")) %>%
  distinct(sample)

All.ASVs.with.tax %>%
  group_by (Blast.Species, sample) %>%
  summarise (by.spp.blast = sum(nReads)) %>%
  arrange(desc(by.spp.blast)) -> summary.by.spp.blast

summary.by.spp.blast %>%
  filter(str_detect( string = Blast.Species, pattern = "No"))

All.ASVs.with.tax %>%
  group_by (species.insect) %>%
  summarise (by.spp.insect = sum(nReads)) %>%
  arrange(desc(by.spp.insect)) %>% 
  full_join(summary.by.spp.blast, by = c("species.insect" = "Blast.Species" ))

```

```{r}

All.ASVs.with.tax %>%
  group_by(rank) %>%
  summarise (n = sum(nReads)) %>%
  ggplot (aes(x = fct_reorder(rank, desc(n)), y = n)) +
  geom_col() +
  labs (x = "Rank")

#### Generate a table of the number of reads identified at least to family level

ALL.ASVs %>% left_join(all.hashes.blast) %>%
  group_by(Family) %>%
  summarise (n = sum(nReads)) %>%
  ggplot (aes(x = fct_reorder(Family, desc(n)), y = n)) +
  geom_col() +
  labs (x = "Family") +
  theme (axis.text.x = element_blank(),
         axis.ticks.x = element_blank())

All.ASVs.with.tax %>% 
  separate(sample, into = c("SampleShort", "rep"), sep = "\\.", remove = F) -> All.ASVs.with.tax

summary(All.ASVs.with.tax$SampleShort %in% OA_Samples$SampleShort)

All.ASVs.with.tax %>% filter (!SampleShort %in% OA_Samples$SampleShort)

All.ASVs.with.tax %>% filter (phylum != "" ) %>%
  mutate (family = case_when(family == "" ~ taxon,
                             TRUE         ~ family)) %>% 
  separate(sample, into = c("SampleShort", "rep"), sep = "\\.", remove = F) -> All.ASVs.with.tax

All.ASVs.with.tax %>% inner_join(OA_Samples, by = "SampleShort")  %>% 
  group_by(phylum, family, Site) %>%
  dplyr::summarise(nReads = sum(nReads)) %>%
  ggplot(aes(x= phylum, y = nReads)) +
  geom_col()

All.ASVs.with.tax %>% 
  group_by(sample, SampleShort, family) %>%
  summarise(nReads = sum(nReads))  %>%
  inner_join(OA_Samples, by = "SampleShort") -> ASVs.by.family

  summarise(nReads = sum(nReads))

```

Now we can use this dataset to elaborate our first exploratory analysis

## How many families are shared by each level of replication

```{r shared families}

ASVs.by.family %>% separate(sample, into = c("biol", "rep"), sep = "\\.", remove = F) %>%
  separate(biol, into = c("Site", "date"), remove = F) %>% 
  separate(date, into = "date", sep = -1) %>%
  separate (biol, into = "Site-date", sep = -1, remove = F) -> ASVs.by.family

## Now Start with site

ASVs.by.family %>% group_by(family) %>%
  summarise(n_sites = n_distinct (Site))

```